---
title: "GIS Homework 4"
output: html_document
---
```{r}
library(sf)
library(here)
library(janitor)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(rmapshaper)
library(maptools)
library(RColorBrewer)
library(classInt)
library(sp)
library(rgeos)
library(tmap)
library(tmaptools)
library(rgdal)
library(geojsonio)
library(readr)

```

read in global gender inequality data and world countries shapefile
"HRD21-22_composite_indices.csv" obtainable at: https://hdr.undp.org/sites/default/files/2021-22_HDR/HDR21-22_Composite_indices_complete_time_series.csv
"World_Countries_T.shp" obtainable in raw form at: https://hub.arcgis.com/datasets/2b93b06dc0dc4e809d3c8db5cb96ba69_0/explore?location=-2.688200%2C0.000000%2C1.41 
But it has been manually edited Turkiye to Turkey for a new shapefile World_Countries_T (elaborated on in Step 3a)
```{r}
raw_inequality_index <- read_csv(here("hw4_data", 
                                      "HDR21-22_composite_indices.csv"))

world <- st_read(here("hw4_data", 
                      "World_Countries_T.shp"))
```
#Task 
Join the global gender inequality index to spatial data of the World, creating a new column of difference in inequality between 2010 and 2019

We have global gender inequality index data from 1990-2021 in the columns labelled 
"gii_1990" to "gii_2021" 
Steps: 
1a. Extract "iso3" and "Country" columns needed for spatial join
1b. Extract "gii_2010" and "gii_2021" columns, and maybe_gii_rank_2021 for deeper analysis. 
2. Create a new column that computes difference between gii_2010 and gii_2021.
3a. Create new column "iso3c" for accurate joining of spatial data (world_cleaned) to gii_data
3b. Spatial join to world countries shapefile
4. Map difference

### 1. Extract necessary columns from gii data
```{r}

gii_data <- raw_inequality_index[, c("iso3", "country", 
                         "gii_rank_2021", "gii_2010", "gii_2021")]

```
### 2. Compute difference between gii_2010 and gii_2021

```{r}
gii_data <- gii_data %>%
  mutate(gii_diff=(gii_2021 - gii_2010))
```
### 3a. 

```{r}
#try using countrycode package to convert country names in "world" spatial data to iso3 code
library(countrycode)
world_cleaned <- world %>%
  mutate(iso3c = 
           countrycode(world[[2]], origin = "country.name", destination = "iso3c")
         )
```
In preliminary running of code, countrycode returned the following message: 
"Some values were not matched unambiguously: Azores, Bonaire, Canarias, Glorioso Islands, Juan De Nova Island, Madeira, Micronesia, Saba, Saint Eustatius, Saint Martin, Turkiye"
Manually edited Turkiye to Turkey for a new shapefile World_Countries_T
The other values are negligible and likely to not have gii data.
New message should read: 
"Some values were not matched unambiguously: Azores, Bonaire, Canarias, Glorioso Islands, Juan De Nova Island, Madeira, Micronesia, Réunion, Saba, Saint Eustatius, Saint Martin"
Moving on...
### 3b. Spatial join gii_data to wold countries shapefile

```{r}
joined_data <- world_cleaned %>% 
  left_join(., 
            gii_data,
            by = c("iso3c" = "iso3"))
```
### 4. Map the difference between gii data in 2010 and 2021

# Change1:  change the color from the minimum value with dark red to dark green, since it means that they improve the most to have lower gender inequality

# Change2: add country name text for those countries that have a strong improvement on gender inequality

```{r}
tm_shape(joined_data) + 
  tm_polygons(col = "gii_diff", 
              midpoint = NA,
              palette = '-YlGn') +
  tm_layout(main.title = "Gender Inequality from 2010-2021", 
            main.title.position = "center",
            legend.position = c("left", "bottom")) + 
  tm_shape(joined_data %>% filter(gii_diff < -0.1)) + 
  tm_text("iso3c", size = 0.4)
```

Feel free to adapt the code to your needs! 



